Quantifying Infants' Everyday Restrained Experiences in the Home Using Wearable Inertial Sensors

### 2025-11-17 manuscript_restrained_classification validation.qmd
fixed typo

abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives. However, previous survey-based and video-based methods cannot simultaneously provide continuous, full-day accounts of infant restraint experiences. This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensors movement data derived from an existing data set. The data set includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months. We annotated infants’ restrained states in the first 1.5-hour video recordings of each session as ground truth labels. The supervised machine-learning model achieved high accuracy (89%) and substantial kappa agreement (κ = .73) compared with human-coded ground truth. The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length. The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant overall restraint time throughout the day. In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."

@@

Physical restraint is a common feature in infants’ everyday lives. Infants are routinely held or carried by caregivers, and infants are often restrained in devices such as highchairs, walkers, and strollers [@Birken2015s]. As an externally regulated aspect of infants’ daily life, restraint shapes infants’ experiences based on caregivers’ decisions--caregivers, not infants, choose when infants are restrained. Restraint limits infants’ physical activities and might hinder infants’ motor development [e.g., @Pin2007z]. Restraint also alters experiences that are relevant for perceptual and cognitive development by determining what is accessible in the physical world. For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount and quality of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h]. Therefore, characterizing infants’ restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.

@@

In addition to measuring the total duration of infants’ restrained time, prior research has also examined associations between restraint and developmental outcomes. Motor development appears to be the domain most directly associated with restraint. Various studies have shown that restraint is related to less proficiency in infant motor skills and temporary delay in milestone achievement [@Abbott2001u; @Bartlett2003w; @Carson2022n; @Karasik2023d; @Pin2007z]. Physiological studies tracking infants’ body movements (e.g., spine, legs, hips) in real time suggest that reduced movement opportunities may underlie these effects [@Jiang2016z; @Siddicky2020f; @Siddicky2021c]. Restraint may also influence other aspects of infant development by shaping infants’ learning opportunities in real time. Infants produce fewer object interactions when restrained compared to when they are unrestrained [@Franchak2024l]. Moreover, infants hear less amount of and less consistent adult language when restrained compared to unrestrained [@Malachowski2023a].

@@

Video recording and human annotation overcomes the above drawbacks of survey-based methods, but video is subject to other limitations. One disadvantage of video observation is being obtrusive. The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors. Another limitation is that video recording and video annotation are both labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies. Indeed, prior work that videotaped and manually annotated infants’ restraint had short durations of recording (e.g., half an hour in @Springfield2025p, and one hour in @Wang2025v). Within the short duration of recording, researchers tend to record the period when infants are awake and active. In contrast, periods when the infants are restrained, like seated in front of TV or in highchairs, would likely be skipped. Thus, video recording is subject to sampling bias and might not be representative of infants’ whole-day experiences.

@@

A system that uses wearable sensors (i.e., inertial movement units) to record real-time movement data and machine learning models to classify movement categories can overcome the limitations listed above. First, by collecting infants’ movement data through sensors, the system can objectively capture infants’ motion without relying on a respondent or a video annotator’s interpretation. Second, the sensors can provide moment-to-moment observation that is continuous and rich in time-sequence information. Third, sensors embedded in baby garments are mobile and unobtrusive—an experimenter does not need to follow the infant with a camera to record movements. Fourth, the system can record and process full-day or multiple-day data at a time due to the sensors’ sufficient battery life and automated classification of machine learning models.

@@

Detecting restrained periods might be more challenging than classifying body postures. First, infants may show similar body movements whether restrained or not, making restrained and unrestrained classes more difficult to differentiate. For example, infants might kick legs when they are lying without restraint and also when they are strapped in a highchair. Possible distinctions might be manifested in reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect. Second, there are lots of possible restraint forms infants can present--a caregiver can hold an infant up straight or in arms while breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d]. Accordingly, the wearable sensor system should be robust enough to generalize across heterogeneous restraint scenarios. Thus, the dual requirement of sensitivity and robustness is the key challenge for developing a reliable model to detect infant restraint in the home.

@@

The data set is hosted on Databrary (Franchak, 2023). The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged 4 to 7 months, and `r apa_num(age[2,2])` older infants aged 11 to 14 months. Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study for only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions. Thirty-four of the infants are female and thirty-one are male. Caregivers reported the infants’ race as White (*n* = 23), Hispanic or Latinx (*n* = 8), African American (*n* = 2), Asian (*n* = 2), more than one (*n* = 15), or not fitting into the above races or chose not to answer (*n* = 15).

@@

In the data set, a special infant garment was customized to secure four lightweight wearable sensors (MC10 Biostamp) at four specific locations (i.e., right thigh, left thigh, right ankle, and left ankle). Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations). The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for whole-day-long recording. An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record synchronized third-person video of the first 1.5 hours into the session as ground truth.

@@

Using the 1.5-hour videos, our team annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/). Restrained and unrestrained was defined as a binary category, meaning that all times were coded as one of the two categories. A period was considered missing if the infant went out of view; missing periods were excluded from further analysis. Restraint was defined as when infants could not initiate or control their own movements, which included being held or carried by caregivers, or being restrained in devices that constrain them from changing body position. Unrestrained periods, in contrast, were when infants could freely change their body position or locomote. For example, an infant was considered restrained when strapped into a high chair seat, but not when they were placed in a crib and free to change position. Each video file was coded by two coders: a primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of each video. Inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label. The human annotation reached a high agreement with an overall accuracy of `r apa_num(stats_human[1,1]*100)`%, ranging from `r apa_num(stats_human[1,2]*100)`% to `r apa_num(stats_human[1,3]*100)`%, and an average kappa of `r apa_num(stats_human[1,4])`, ranging from `r apa_num(stats_human[1,5])` to `r apa_num(stats_human[1,6])`.

@@

The third step was to train and validate supervised machine learning models that classify the motion features into restrained status. We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia. We used a leave-one-out cross-validation approach, where we held out one session at each iteration as testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data. After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.

@@

```{r fig-prevalencegap}
#| fig-cap: Agreement between model-predicted and human annotated prevalence of restraint
#| fig-width: 8
#| fig-height: 4
#| apa-note: Prevalence is defined as the proportion of windows predicted (or annotated) as restrained among all windows. (A) shows the correlation of prevalence between model prediction and human annotation. The black dashed identity line indicates perfect agreement. Each circular symbol represents one session for a specific window length (denoted by color). (B) shows the difference of prevalence between human annotation and model prediction (i.e., prediction error). A negative prediction error indicates the model estimated less amount of time infants were restrained compared to human annotation. The horizontal line in the box denotes the median; the boarder lines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).

@@

@fig-prevalencegap A shows that, at the session level, all window lengths provided reliable estimate of restrained time. We defined prevalence as the proportion of windows classified as restrained among all windows for each session. The prevalence predicted by the model was highly correlated with the prevalence annotated by human coders (*r* = `r apa_num(table_prev[1,4])` for 4 s, *r* = `r apa_num(table_prev[2,4])` for 16 s, and *r* = `r apa_num(table_prev[3,4])` for 30 s). Thus, model predictions are suitable for capturing individual differences in restrained time between infants.

### Models were biased towards predicted unrestrained

Despite the overall strong accuracy, models using all three window lengths were biased towards predicting that infants were unrestrained rather than restrained. @fig-cibtwctg shows the difference of a sensitivity and PPV between restrained class and unrestrained class within a single window length. Larger difference scores (away from 0) indicate larger asymmetry in predicting one category rather than the other. @fig-cibtwctg A shows that the sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI \[`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`\], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI \[`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI \[`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`\]. The negative difference scores for sensitivity across all window lengths indicate that the model was more sensitive in detecting the unrestrained class compared to restrained class regardless of window length. @fig-cibtwctg B shows that the PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI \[`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`\], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI \[`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI \[`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`\]. The positive differences of PPV across all window lengths indicate that the model’s prediction of restrained periods was more credible compared to its prediction of unrestrained periods. Notably, while none of the window lengths were free from bias, the 30 s window presented smaller offsets of metric values between restrained and unrestrained detection.

@@

We applied the best model based on the analyses in the previous (30-s window length) to create full-day predictions for all sessions that included full-day recording. We excluded eleven sessions whose recording time was less than 3 hours per session, which were useful for training and validating the models but insufficient for measuring day-long experience. We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal. The remaining 129 sessions had an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).

@@

Why could a longer window length (i.e., 30 s) provide more balanced detection? One explanation is that a longer window length can better capture the contextual information of restraint. Consider the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not for most of the time, but one subtle occurrence of distinctive movement can signal restraint. For example, an infant can be sitting on the floor unrestrained, on caregiver’s lap restrained, or strapped in a seat restrained. It is hard to distinguish if the infant is restrained while sitting still, but when the infant moves their limbs, restraint may present differences, such as a stronger damping in the motion decay because the support (i.e., caregiver’s laps, or seats) absorbs the motion. These differences only present themselves in the episodes of movement, not throughout the whole period. A longer window length has a better chance to capture the episodes. Another explanation is that a longer window length can improve generalization across a variety of restraint forms. Infants can be restrained in various ways—such as in stationary seats, in jumpers, in swings, or in caregivers’ arms, either still or while being carried around. Different restraint forms are presented by different motion features. By encompassing a broader range of movement patterns, a longer window can provide more informative and generalizable features for the model to account for variability within the restrained class. Given a restrained bout usually lasts minutes rather than seconds, a 30 s window length presents a good balance.

@@

Beyond expanding the range of classifiable behaviors, future works should also broaden the scenarios in which wearable sensors are applied. To date, most applications of wearable sensors have been limited to laboratory or in-home environments. Infants’ outdoor experiences remain largely underexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration [@Schneider2025d; @Herrington1998o; @Fjortoft2000o]. How much time do infants spend restrained in strollers or car seats and while worn by parents in wearable carriers? Which positions are they frequently in when restrained out of the house? Validating sensor-based models in outdoor scenarios can open new avenues for investigation.

@@

## Advances in Understanding Full-day Real-time Restraint

@@

Having a full-day continuous profile of infant restraint is also essential in understanding infant real-time sensorimotor experiences. An overall estimate of restraint time is not enough to examine the mechanisms between restraint practices and development outcomes. Take motor development as an example. @Karasik2023d found that Tajik infants fell behind US infants in achieving motor milestones despite the comparable overall restrained time between cultural groups. It is likely due to the prolonged individual restrained bouts in Tajik infants [@Karasik2022a]. The current system allows researchers to characterize the temporal structure of restraint and examine the relationship between restraint and other experiences, such as language, object interaction, and social interaction. A prior study [@Malachowski2023a] found device restraint is associated with reduced adult language input using ecological momentary assessment. However, a continuous, full-day recording of restraint and speech can capture the temporal distribution rather than snapshots. Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout? How do specific language inputs (e.g., referential words) distribute throughout a day based on restraint states and restraint types? Additionally, real-time restraint can also be linked to infants’ self-initiated explorations. For example, how do infants’ own vocalizations, attention, locomotion, or object interactions differ across restrained and unrestrained? In short, continuous, full-day quantification of infant restraint offers a foundation for linking caregiver practices to infants’ moment-to-moment sensorimotor experiences, thereby shaping infants' developmental outcomes.

>>>>

abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives. However, previous survey-based and video-based methods cannot simultaneously provide continuous, full-day accounts of infant restraint experiences. This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensor movement data derived from an existing data set. The data set includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months. We annotated infants’ restrained states in the first 1.5-hour video recordings of each session as ground truth labels. The supervised machine-learning model achieved high accuracy (89%) and substantial kappa agreement (κ = .73) compared with human-coded ground truth. The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length. The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infants' overall restraint time throughout the day. In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."

@@

Physical restraint is a common feature in infants’ everyday lives. Infants are routinely held or carried by caregivers, and infants are often restrained in devices such as highchairs, walkers, and strollers [@Birken2015s]. As an externally regulated aspect of infants’ daily life, restraint shapes infants’ experiences based on caregivers’ decisions--caregivers, not infants, choose when infants are restrained. Restraint limits infants’ physical activities and might hinder infants’ motor development [e.g., @Pin2007z]. Restraint also alters experiences that are relevant for perceptual and cognitive development by determining what is accessible in the physical world. For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h]. Therefore, characterizing infants’ restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.

@@

In addition to measuring the total duration of infants’ restrained time, prior research has also examined associations between restraint and developmental outcomes. Motor development appears to be the domain most directly associated with restraint. Various studies have shown that restraint is related to less proficiency in infant motor skills and temporary delay in milestone achievement [@Abbott2001u; @Bartlett2003w; @Carson2022n; @Karasik2023d; @Pin2007z]. Physiological studies tracking infants’ body movements (e.g., spine, legs, hips) in real time suggest that reduced movement opportunities may underlie these effects [@Jiang2016z; @Siddicky2020f; @Siddicky2021c]. Restraint may also influence other aspects of infant development by shaping infants’ learning opportunities in real time. Infants produce fewer object interactions when restrained compared to when they are unrestrained [@Franchak2024l]. Moreover, infants hear less frequent and less consistent adult language when restrained compared to unrestrained [@Malachowski2023a].

@@

Video recording and human annotation overcome the above drawbacks of survey-based methods, but video is subject to other limitations. One disadvantage of video observation is being obtrusive. The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors. Another limitation is that video recording and video annotation are both labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies. Indeed, prior work that videotaped and manually annotated infants’ restraint had short durations of recording (e.g., half an hour in @Springfield2025p, and one hour in @Wang2025v). Within the short duration of recording, researchers tend to record the period when infants are awake and active. In contrast, periods when the infants are restrained, like seated in front of the TV or seated in highchairs, would likely be skipped. Thus, video recording is subject to sampling bias and might not be representative of infants’ whole-day experiences.

@@

A system that uses wearable sensors (i.e., inertial movement units) to record real-time movement data and machine learning models to classify movement categories can overcome the limitations listed above. First, by collecting infants’ movement data through sensors, the system can objectively capture infants’ motion without relying on a respondent or a video annotator’s interpretation. Second, the sensors can provide moment-to-moment observation that is continuous and rich in time-sequence information. Third, sensors embedded in baby garments are mobile and unobtrusive—an experimenter does not need to follow the infant with a camera to record movements. Fourth, the system can record and process full-day and multiple-day data due to the sensors’ sufficient battery life and automated classification of machine learning models.

@@

Detecting restrained periods might be more challenging than classifying body postures. First, infants may show similar body movements whether restrained or not, making restrained and unrestrained classes more difficult to differentiate. For example, infants might kick legs when they are lying without restraint and also when they are strapped in a highchair. Possible distinctions might be manifested in a reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect. Second, there are lots of possible restraint forms infants can present--a caregiver can hold an infant up straight or in arms while breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d]. Accordingly, the wearable sensor system should be robust enough to generalize across heterogeneous restraint scenarios. Thus, the dual requirement of sensitivity and robustness is the key challenge for developing a reliable model to detect infant restraint in the home.

@@

The data set is hosted on Databrary (Franchak, 2023). The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged 4 to 7 months, and `r apa_num(age[2,2])` older infants aged 11 to 14 months. Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions. Thirty-four of the infants are female and thirty-one are male. Caregivers reported the infants’ race as White (*n* = 23), Hispanic or Latinx (*n* = 8), African American (*n* = 2), Asian (*n* = 2), more than one (*n* = 15), or not fitting into the above races or chose not to answer (*n* = 15).

@@

In the data set, a special infant garment was customized to secure four lightweight wearable sensors (MC10 Biostamp) at four specific locations (i.e., right thigh, left thigh, right ankle, and left ankle). Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations). The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for daylong recording. An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record synchronized third-person video of the first 1.5 hours into the session as ground truth.

@@

Using the 1.5-hour videos, our team annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/). Restrained and unrestrained were defined as a binary category, meaning that all times were coded as one of the two categories. A period was considered missing if the infant went out of view; missing periods were excluded from further analysis. Restraint was defined as when infants could not initiate or control their own movements, which included being held or carried by caregivers, or being restrained in devices that constrain them from changing body position. Unrestrained periods, in contrast, were when infants could freely change their body position or locomote. For example, an infant was considered restrained when strapped into a high chair seat, but not when they were placed in a crib and free to change position. Each video file was coded by two coders: a primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of each video. Inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label. The human annotation reached a high agreement with an overall accuracy of `r apa_num(stats_human[1,1]*100)`%, ranging from `r apa_num(stats_human[1,2]*100)`% to `r apa_num(stats_human[1,3]*100)`%, and an average kappa of `r apa_num(stats_human[1,4])`, ranging from `r apa_num(stats_human[1,5])` to `r apa_num(stats_human[1,6])`.

@@

The third step was to train and validate supervised machine learning models that classify the motion features into restrained status. We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia. We used a leave-one-out cross-validation approach, where we held out one session at each iteration as a testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data. After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.

@@

```{r fig-prevalencegap}
#| fig-cap: Agreement between model-predicted and human annotated prevalence of restraint
#| fig-width: 8
#| fig-height: 4
#| apa-note: Prevalence is defined as the proportion of windows predicted (or annotated) as restrained among all windows. (A) shows the correlation of prevalence between model prediction and human annotation. The black dashed identity line indicates perfect agreement. Each circular symbol represents one session for a specific window length (denoted by color). (B) shows the difference of prevalence between human annotation and model prediction (i.e., prediction error). A negative prediction error indicates the model estimated less amount of time infants were restrained compared to human annotation. The horizontal line in the box denotes the median; the borderlines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).

@@

@fig-prevalencegap A shows that, at the session level, all window lengths provided reliable estimates of restrained time. We defined prevalence as the proportion of windows classified as restrained among all windows for each session. The prevalence predicted by the model was highly correlated with the prevalence annotated by human coders (*r* = `r apa_num(table_prev[1,4])` for 4 s, *r* = `r apa_num(table_prev[2,4])` for 16 s, and *r* = `r apa_num(table_prev[3,4])` for 30 s). Thus, model predictions are suitable for capturing individual differences in restrained time between infants.

### Models Were Biased Toward Predicted Unrestrained

Despite the overall strong accuracy, models using all three window lengths were biased towards predicting that infants were unrestrained rather than restrained. @fig-cibtwctg shows the difference of sensitivity and PPV between restrained class and unrestrained class within a single window length. Larger difference scores (away from 0) indicate larger asymmetry in predicting one category rather than the other. @fig-cibtwctg A shows that the sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI \[`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`\], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI \[`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI \[`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`\]. The negative difference scores for sensitivity across all window lengths indicate that the model was more sensitive in detecting the unrestrained class compared to restrained class regardless of window length. @fig-cibtwctg B shows that the PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI \[`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`\], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI \[`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI \[`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`\]. The positive differences of PPV across all window lengths indicate that the model’s prediction of restrained periods was more credible compared to its prediction of unrestrained periods. Notably, while none of the window lengths were free from bias, the 30 s window presented smaller offsets of metric values between restrained and unrestrained detection.

@@

We applied the best model based on the analyses in the previous section (30-s window length) to create full-day predictions for all sessions that included full-day recording. We excluded eleven sessions whose recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience. We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal. The remaining 129 sessions had an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).

@@

Why could a longer window length (i.e., 30 s) provide more balanced detection? One explanation is that a longer window length can better capture the contextual information of restraint. Consider the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not for most of the time, but one subtle occurrence of distinctive movement can signal restraint. For example, an infant can be sitting on the floor unrestrained, on a caregiver’s lap restrained, or strapped in a seat restrained. It is hard to distinguish if the infant is restrained while sitting still, but when the infant moves their limbs, restraint may present differences, such as a stronger damping in the motion decay because the support (i.e., caregiver’s laps, or seats) absorbs the motion. These differences only present themselves in the episodes of movement, not throughout the whole period. A longer window length has a better chance to capture the episodes. Another explanation is that a longer window length can improve generalization across a variety of restraint forms. Infants can be restrained in various ways—such as in stationary seats, jumpers, swings, or caregivers’ arms, either still or while being carried around. Different restraint forms are presented by different motion features. By encompassing a broader range of movement patterns, a longer window can provide more informative and generalizable features for the model to account for variability within the restrained class. Given a restrained bout usually lasts minutes rather than seconds, a 30 s window length presents a good balance.

@@

Beyond expanding the range of classifiable behaviors, future works should also broaden the scenarios in which wearable sensors are applied. To date, most applications of wearable sensors have been limited to laboratory or in-home environments. Infants’ outdoor experiences remain largely underexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration [@Schneider2025d; @Herrington1998o; @Fjortoft2000o]. How much time do infants spend restrained in strollers, car seats, or in carriers worn by caregivers? Which positions are they frequently in when restrained out of the house? Validating sensor-based models in outdoor scenarios can open new avenues for investigation.

@@

## Advances in Understanding Full-Day Real-Time Restraint

@@

Having a full-day continuous profile of infant restraint is also essential in understanding infant real-time sensorimotor experiences. An overall estimate of restraint time is not enough to examine the mechanisms between restraint practices and developmental outcomes. Take motor development as an example. @Karasik2023d found that Tajik infants fell behind US infants in achieving motor milestones despite the comparable overall restrained time between cultural groups. It is likely due to the prolonged individual restrained bouts in Tajik infants [@Karasik2022a]. The current system allows researchers to characterize the temporal structure of restraint and examine the relationship between restraint and other experiences, such as language, object interaction, and social interaction. A prior study [@Malachowski2023a] found device restraint is associated with reduced adult language input using ecological momentary assessment. However, a continuous, full-day recording of restraint and speech can capture the temporal distribution rather than snapshots. Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout? How are specific language inputs (e.g., referential words) distributed throughout a day based on restraint states and restraint types? Additionally, real-time restraint can also be linked to infants’ self-initiated explorations. For example, how do infants’ own vocalizations, attention, locomotion, or object interactions differ across restrained and unrestrained? In short, continuous, full-day quantification of infant restraint offers a foundation for linking caregiver practices to infants’ moment-to-moment sensorimotor experiences, thereby shaping infants' developmental outcomes.
